Individual heterogeneity in functional topography of the default-mode network in medial prefrontal cortex

### 2019-03-27 Manuscript.Rmd
Added papers to bibliography, and finished revising the intro

  A useful way to characterize individual-specific brain organization is to examine patterns of resting-state functional connectivity (rsFC; i.e. time-series correlations between pairs of locations in the brain). Connectome-based analyses of rsFC have been fruitful in characterizing individualized functional topologies that correspond well to task-induced activity [@Gordon2017; @Laumann2015; @Tobyne2017; @Smith2009]. A popular way to analyze these connectomes is to rely on graph theoretic methods, which are properly equiped to understand brain network dynamics [@Bassett2018a]. In the context of network analysis, community detection algorithms have been successfully used to subdivide brain networks into multiple distinct communities [@Garcia2017]. Such communities represent clusters of network nodes that are more connected with each other than with the rest of the network [@Fortunato2016]. Among these algorithms, spectral partitioning (SP) efficiently captures  [@Fiedler1975; @Belkin2003]. SP has recently been used to characterize the posterior-anterior functional gradient of the insula using rsfMRI data [@Tian2018]. This method can thus capture a refined group of brain regions that more cohesively share activation patterns during rest.

  In this study, we aimed to subsection regions commonly attributed to both DMN and subjective value into subject-specific DMN and non-DMN partitions, quantifying the degree of topographical heterogeneity present in this network within and across individuals. We did this by capitalizing on the respective strengths of meta-analytic and subject-specific analyses of brain networks. First, we established a search space by selecting parcels from an established brain atlas [@Glasser2016] that considerably overlapped with previously-defined DMN and limbic networks along the medial wall [@Yeo2011]. A meta-analysis of DMN and SV literatures revealed areas in PCC and mPFC that were uniquely related to DMN and SV, respectively (these areas were used as markers to identify the affiliation of each individual's communities). We then produced rsFC networks of all the surface vertices within these regions of interest (ROI) for each of 100 individual resting state dataset from the Human Connectome Project (HCP)[@VanEssen2012], and estimated the extent of DMN coverage through spectral partitioning. Focusing on individual vertices in this search space rather than the parcels (as is usual in brain network analyses) allowed us to finely delineate the topographical extent of each community. We found that these areas of overlap can be partitioned into two communities that topographically vary across subjects, but that follow a comon organizational principle. This partition was in substantial agreement with results from a different community detection algorithm (modularity). A sliding window analysis showed that spectral partitioning allows for a more immediate identification of noisy viertices with unreliable affiliation. Refining the partitions by filtering out these nodes resulted in idiosyncratic partitions that were highly stable across scanning runs per individual, and this pattern was higher for mPFC. We show that partitions produced by spectral partitioning had higher test-retest reliability per subject than common seed-based functional connectivity, without the need of group-drived priors [@Kong2018]. Lastly, we provide evidence that these communities resembled an organizational scheme recently observed in a handful of subjects [@Braga2017]. Our work highlights the usefulness of estimating brain effects at the individual level in these regions, and points to limitations of aggregative methods in determining the degree of neural overlap between DMN and valuation.

@@

# Store thre FV thresholds for all subject 
threshDF <- data_frame(SubjID = substring(subjList,1,6),
                       Lower = NA,
                       Upper = NA)
for (i in seq_along(tempList)) {
  if (is.null(tempList[[i]]$result)) {
    # if there wass no clear threshold (i.e. noisy participants), then leave it as 0
    threshDF[i,2:3] <- c(0,0)
  } else {
    threshDF[i,2] <- as.numeric(tempList[[i]]$result[1])
    threshDF[i,3] <- as.numeric(tempList[[i]]$result[2])
  }
}

@@

### *Partition Comparison: Thresholded Full Datasets*
``` {r Partition comparison: all thresholded, echo=FALSE, message = FALSE, cache = TRUE}
# Apply threshold to Summaries
# IMPORTANT: THIS PROCESS OVERWRITES THE VARIABLES FROM THE FIRST COMPARISON SECTION
# THIS IS DONE FOR STORAGE-SAKE, BUT IT MIGHT BE WORTH GIVING THESE TEIR OWN NAME
if (threshMaps) {
  for (i in seq(nSubj)) {
  Summaries[[i]] <- Summaries[[i]] %>%
    mutate(FiedlerVec = ifelse(FiedlerVec > as.numeric(threshDF[i,2]) & FiedlerVec < as.numeric(threshDF[i,3]), 0, FiedlerVec)) %>%
    mutate(FiedlerBinary = ifelse(FiedlerVec > 0, 1, ifelse(FiedlerVec < 0, 0, 0.5)))
  }
}

>>>>

  A useful way to characterize individual-specific brain organization is to examine patterns of resting-state functional connectivity (rsFC; i.e. time-series correlations between pairs of locations in the brain). Connectome-based analyses of rsFC have been fruitful in characterizing individualized functional topologies that correspond well to task-induced activity [@Gordon2017; @Laumann2015; @Tobyne2017; @Smith2009]. A popular way to analyze these connectomes is to rely on graph theoretic methods, which are properly equiped to understand brain network dynamics [@Bassett2018a]. In the context of network analysis, community detection algorithms have been successfully used to subdivide brain networks into multiple distinct communities [@Garcia2017]. Such communities represent clusters of network nodes that are more connected with each other than with the rest of the network [@Fortunato2016]. In attempting to refine the DMN topography at the individual level, we focus on dividing mPFC and PCC into communities encapsulating what belongs and does not belong to DMN. Spectral partitioning (SP) is an efficient community detection algorithm that deterministically forces such bisections [@Fiedler1975; @Belkin2003]. For a network composed of pairwise correlations among its nodes, SP partitions the network by binarizing the eigenvector linked to the second-to-lowest eigenvalue of the adjusted (Laplacian) correlation matrix according to its sign (i.e. positive or negative number). SP has recently been used to characterize the posterior-anterior functional gradient of the insula using rsfMRI data [@Tian2018], and was recently shown to robustly and reliably separate both simulated and actual primate ECoG networks [@Toker2019]. This method can therefore capture a refined group of brain regions that more cohesively share activation patterns during rest.

  In this study, we aimed to subdivide regions commonly attributed to both DMN and subjective valuation into individual-specific DMN and non-DMN communities, quantifying the degree of topographical heterogeneity present in this network within and across individuals. We did this by capitalizing on the respective strengths of meta-analytic and subject-specific analyses of brain networks. First, we established a search space by selecting parcels from an established brain atlas [@Glasser2016] that corresponded to previously defined DMN and limbic networks along the medial cortical wall [@Yeo2011]. A cortical surface-based meta-analysis of DMN and SV literatures identified a parcel in PCC that was DMN-specific at the aggregate level. We then derived an rsFC network of all the surface vertices within the search space for each of 100 individual resting-state fMRI data sets from the Human Connectome Project (HCP)[@VanEssen2012], and used the SP algorithm to subdivide each individual's network into DMN and non-DMN communities (defined according to the degree of coverage of the PCC parcel). Focusing on individual vertices in the search space rather than the parcels (as is typical in brain network analyses) allowed us to finely delineate the topographical extent of each community. The resulting communities varied topographically across subjects, while also appearing to follow common organizational principles. A combination of sliding window and test-retest analyses showed that these partitionings were highly similar across scanning days within (but not between) individuals, and that this idiosyncrasy was higher in mPFC. Partitionings obtained from the SP algorithm had higher test-retest reliability than did analogous results from seed-based functional connectivity. Lastly, we provide evidence that the resulting communities resembled an organizational scheme recently observed in a handful of subjects [@Braga2017]. Our work highlights the usefulness of estimating brain effects at the individual level in mPFC and PCC, and provides a new framework and tool set for future investigations of overlap across cognitive domains.

@@

# Store thre FV thresholds for all subject 
threshDF <- data_frame(SubjID = substring(subjList, 1, 6),
                       Lower = NA,
                       Upper = NA)
for (i in seq_along(tempList)) {
  if (is.null(tempList[[i]]$result)) {
    # if there wass no clear threshold (i.e. noisy participants), then leave it as 0
    threshDF[i, 2:3] <- c(0, 0)
  } else {
    threshDF[i, 2] <- as.numeric(tempList[[i]]$result[1])
    threshDF[i, 3] <- as.numeric(tempList[[i]]$result[2])
  }
}

@@

### *Partition Comparison: Thresholded Full Datasets*
``` {r Partition comparison: all thresholded, echo=FALSE, message = FALSE, cache = TRUE}
# Apply threshold to Summaries
# IMPORTANT: THIS PROCESS OVERWRITES THE VARIABLES FROM THE FIRST COMPARISON SECTION
# THIS IS DONE FOR STORAGE-SAKE, BUT IT MIGHT BE WORTH GIVING THESE TEIR OWN NAME
if (threshMaps) {
  for (i in seq(nSubj)) {
  Summaries[[i]] <- Summaries[[i]] %>%
    mutate(FiedlerVec = ifelse(FiedlerVec > as.numeric(threshDF[i, 2]) & FiedlerVec < as.numeric(threshDF[i, 3]), 0, FiedlerVec)) %>%
    mutate(FiedlerBinary = ifelse(FiedlerVec > 0, 1, ifelse(FiedlerVec < 0, 0, 0.5)))
  }
}
